---
title: "Singapore Birth Rate EDA"
author: "Kiran Shankaran"
date: "2026-03-27"
output: html_document
---

## 1. Introduction

GitHub Repository:
https://github.com/kiranshank12519/Time-Series-Project-Singapore-Births-Forecast/tree/main

This report presents an exploratory data analysis (EDA) of Singapore's Total Live Births (TLB) and Total Fertility Rate (TFR) from 1960 to 2024.The aim is to identify the temporal features of both series, assess stationarity, and propose candidate time series models for the period 1960 to 2012, which will later be used to forecast TLB and TFR from 2013 to 2024.Singapore has experienced a dramatic and sustained decline in fertility since the 1960s, influenced by government population policies,urbanisation,rising education levels, and economic development. These contextual 
factors will be referenced throughout the analysis.

---

## 2. Data Loading and Preparation

The data was sourced from the Singapore Department of Statistics via data.gov.sg. It contains 65 annual observations from 1960 to 2024 for two variables: Total Live Births (TLB) and Total Fertility Rate (TFR). 
No missing values were present in the dataset.
```{r setup, message=FALSE, warning=FALSE}
# Load the data
births <- read.csv("singapore_births.csv")

# Rename columns
colnames(births) <- c("Year", "TFR", "TLB")

# Reverse so years go 1960 to 2024
births <- births[nrow(births):1, ]
rownames(births) <- NULL

# Check it looks right
head(births)
tail(births)

# Convert to time series objects
TLB <- ts(births$TLB, start = 1960, frequency = 1)
TFR <- ts(births$TFR, start = 1960, frequency = 1)

# Training period 1960-2012 only
TLB_train <- window(TLB, start = 1960, end = 2012)
TFR_train <- window(TFR, start = 1960, end = 2012)

```




## 3. Summary Statistics
```{r summary}

# Summary statistics provide a basic numerical description of each series before visual analysis.
# We look at the minimum, maximum, mean and quartiles to understand the overall range and central tendency of TLB and TFR over the full 1960-2024 period.
cat("=== Total Live Births ===\n")
summary(TLB)

cat("\n=== Total Fertility Rate ===\n")
summary(TFR)

```
TLB ranges from a maximum of 61,775 births to a minimum of, with a mean of approximately 45,000 births per year. TFR ranges from 5.76 to a minimum of 0.97, with a mean of 2.20. Both series show a wide range between their minimum and maximum values, which is consistent with a long-run declining trend rather than a stationary series fluctuating around a constant mean.

## 4. Time Series Plots
```{r plot_tlb, fig.width=10, fig.height=5}
# Plotting the full TLB time series from 1960 to 2024.
plot(TLB,
     main = "Total Live Births in Singapore (1960-2024)",
     ylab = "Total Live Births",
     xlab = "Year")
```
```{r plot_tfr, fig.width=10, fig.height=5}
plot(TFR,
     main = "Total Fertility Rate in Singapore (1960-2024)",
     ylab = "Total Fertility Rate (per female)",
     xlab = "Year")
```
Both TLB and TFR show a clear and sustained downward trend over the entire 65-year period with no seasonality present, confirming both series are non-stationary. There are visible changes in the rate of decline at certain points in time and a brief recovery in TLB in the late 1980s, which warrant further investigation.

# Literature analysis
Literature was explored to gain an understanding of the external factors that could potentially explain the features observed in the plots. The acceleration in TLB and TFR decline visible after 1969 is consistent with the introduction of Singapore's Stop-at-Two policy, which penalised families for having more than two children and directly suppressed
birth rates through the 1970s. The brief recovery visible in TLB around 1988-1990 aligns with the government's pro-natalist reversal in 1987, which introduced financial incentives for larger families, though this was insufficient to reverse the long-run downward trajectory.These structural breaks suggest that the underlying drivers of Singapore's fertility decline are not purely demographic. They are shaped by deliberate policy decisions and external shocks that a standard statistical model alone may not fully capture.

## 6. Updated Time Series Plots

The plots are now reproduced with reference lines added at the key policy events identified above, allowing the features of the series to be interpreted more precisely.

```{r plot_tlb_annotated, fig.width=10, fig.height=5}
# Reproduce the TLB plot with policy reference lines added at the years where known government interventions occurred

plot(TLB,
     main = "Total Live Births in Singapore (1960-2024)",
     ylab = "Total Live Births",
     xlab = "Year")
# Adding lines to the time series
abline(v = 1969, col = "red",       lty = 2) 
abline(v = 1987, col = "darkgreen", lty = 2)
abline(v = 2020, col = "orange",    lty = 2)

legend("topright",
       legend = c("Stop-at-Two policy (1969)",
                  "Pro-natalist policy (1987)",
                  "COVID-19 (2020)"),
       col    = c("red", "darkgreen", "orange"),
       lty    = 2,
       cex    = 0.8)
```
```{r plot_tfr_annotated, fig.width=10, fig.height=5}
# Reproduce the TFR plot with policy reference lines and the replacement level line

plot(TFR,
     main = "Total Fertility Rate in Singapore (1960-2024)",
     ylab = "Total Fertility Rate (per female)",
     xlab = "Year")

abline(v = 1969, col = "red",       lty = 2)
abline(v = 1987, col = "darkgreen", lty = 2)
abline(v = 2020, col = "orange",    lty = 2)
abline(h = 2.1,  col = "purple",    lty = 3)

legend("topright",
       legend = c("Stop-at-Two policy (1969)",
                  "Pro-natalist policy (1987)",
                  "COVID-19 (2020)",
                  "Replacement level (2.1)"),
       col    = c("red", "darkgreen", "orange", "purple"),
       lty    = c(2, 2, 2, 3),
       cex    = 0.8)
```
With the reference lines added, the features identified in the raw plots can now be directly linked to their likely causes. The acceleration in TLB and TFR decline after 1969 aligns precisely with the Stop-at-Two policy, the brief recovery in TLB around 1988-1990 corresponds to the pro-natalist reversal, and the sharp dip in 2020 coincides with the COVID-19 pandemic. These annotated plots will serve as the basis for the temporal feature analysis that follows.

# Literature on forecasting births
## 3. Literature Review

To gain an understanding of how others have modelled and forecasted birth rates, a review of recent literature was carried out. He (2023) applied ARIMA models to 74 years of Chinese birth rate data and found ARIMA(0,1,2) to be the best fitting model based on AIC, with first-order differencing sufficient to achieve stationarity. Similarly, a study of Philippine birth data (European Journal of IT and Computer Science, 2025) found that ARIMA-class models outperformed other methods for forecasting declining annual birth rates in a Southeast Asian context directly comparable to Singapore.

From this literature, it can be gathered that ARIMA models with d = 1 are a reasonable starting point for data of this kind, with the AR and MA orders to be identified from the ACF and PACF of the differenced series. However, as noted in the course notes, standard ARIMA models may struggle to capture structural breaks caused by external interventions. Singapore's Stop-at-Two policy (1969) and pro-natalist reversal (1987) represent exactly this kind of disruption, meaning the final model choice may need to be justified on contextual grounds as well as purely statistical ones.

## 5. Temporal Feature Analysis

### 5.1 ACF and PACF of Raw Series
```{r acf_tlb, fig.width=10, fig.height=4}
# The ACF (correlogram) plots the autocorrelation at each lag. As noted in the course notes, if there is a trend the ACF will decrease slowly and remain significant across many lags, indicating non-stationarity.

acf(TLB, lag.max = 20,
    main = "ACF of Total Live Births (1960-2024)")
```
```{r pacf_tlb, fig.width=10, fig.height=4}
# The PACF plots the partial autocorrelation at each lag — the correlation between x_t and x_{t-k} after removing the effect of all intermediate lags.A dominant spike at lag 1 in the PACF suggests an AR(1) component may be present once the trend is removed.

pacf(TLB, lag.max = 20,
     main = "PACF of Total Live Births (1960-2024)")
```
```{r acf_tfr, fig.width=10, fig.height=4}
# ACF of TFR — we expect a similar slowly decaying pattern to TLB given both series share a strong downward trend

acf(TFR, lag.max = 20,
    main = "ACF of Total Fertility Rate (1960-2024)")
```
```{r pacf_tfr, fig.width=10, fig.height=4}
# PACF of TFR to identify any autoregressive structure that may be present in the series

pacf(TFR, lag.max = 20,
     main = "PACF of Total Fertility Rate (1960-2024)")
```

The ACF for both TLB and TFR decreases very slowly and remains significant across many lags which is a signature of a non-stationary series with a strong trend, consistent with the course notes. The PACF for both series shows a dominant spike at lag 1, suggesting that once the trend is removed an autoregressive component may be present. Differencing is required before fitting any ARIMA model.

### 5.2 First-Order Differencing

Since both series have a clear downward trend we apply first-order differencing, defined as:

∇x_t = x_t − x_{t-1}

This removes the trend and is used to determine the differencing order d in our ARIMA(p,d,q) models.
```{r diff_tlb, fig.width=10, fig.height=4}
# Apply first-order differencing to TLB training data using the diff() function and plot the result.A horizontal line at zero is added if the differenced series fluctuates around zero with no clear trend, stationarity has been achieved.

plot(diff(TLB_train),
     main = "First Difference of TLB (1960-2012)",
     ylab = "Change in TLB",
     xlab = "Year",
     col  = "steelblue",
     lwd  = 2)
abline(h = 0, lty = 2)
```
```{r diff_tfr, fig.width=10, fig.height=4}
# Apply first-order differencing to TFR training data and plot the result

plot(diff(TFR_train),
     main = "First Difference of TFR (1960-2012)",
     ylab = "Change in TFR",
     xlab = "Year",
     col  = "darkred",
     lwd  = 2)
abline(h = 0, lty = 2)
```
```{r acf_diff_tlb, fig.width=10, fig.height=4}
# ACF of the differenced TLB series.If differencing has achieved stationarity, the ACF should no longer show the slow decay pattern seen in the raw series. Significant spikes indicate the MA order q to include in the ARIMA model.

acf(diff(TLB_train), lag.max = 20,
    main = "ACF of First Differenced TLB")
```
```{r pacf_diff_tlb, fig.width=10, fig.height=4}
# PACF of the differenced TLB series.Significant spikes in the PACF indicate the AR order p to include in the ARIMA model.

pacf(diff(TLB_train), lag.max = 20,
     main = "PACF of First Differenced TLB")
```
```{r acf_diff_tfr, fig.width=10, fig.height=4}
# ACF of the differenced TFR series

acf(diff(TFR_train), lag.max = 20,
    main = "ACF of First Differenced TFR")
```
```{r pacf_diff_tfr, fig.width=10, fig.height=4}
# PACF of the differenced TFR series

pacf(diff(TFR_train), lag.max = 20,
     main = "PACF of First Differenced TFR")
```

After first-order differencing both series fluctuate around zero with no clear trend, suggesting approximate stationarity has been achieved. This indicates we should set d = 1 in our ARIMA(p,1,q) models. The ACF and PACF of the differenced series are now examined to identify suitable AR and MA orders for model fitting.