---
title: "Singapore Birth Rate EDA"
author: "Kiran Shankaran"
date: "27/03/2026"
output: html_document
---

## 1. Introduction

GitHub Repository:
https://github.com/kiranshank12519/Time-Series-Project-Singapore-Births-Forecast/tree/main

This report presents an exploratory data analysis (EDA) of Singapore's Total Live Births (TLB) and Total Fertility Rate (TFR) from 1960 to 2024.The aim is to identify the temporal features of both series, assess stationarity, and propose candidate time series models for the period 1960 to 2012, which will later be used to forecast TLB and TFR from 2013 to 2024.Singapore has experienced a dramatic and sustained decline in fertility since the 1960s, influenced by government population policies,urbanisation,rising education levels, and economic development. These contextual 
factors will be referenced throughout the analysis.

---

## 2. Data Loading and Preparation

The data was sourced from the Singapore Department of Statistics via data.gov.sg. It contains 65 annual observations from 1960 to 2024 for two variables: Total Live Births (TLB) and Total Fertility Rate (TFR). 
No missing values were present in the dataset.
```{r setup, message=FALSE, warning=FALSE}
# Load the data
births <- read.csv("singapore_births.csv")

# Rename columns
colnames(births) <- c("Year", "TFR", "TLB")

# Reverse so years go 1960 to 2024
births <- births[nrow(births):1, ]
rownames(births) <- NULL

# Check it looks right
head(births)
tail(births)

# Convert to time series objects
TLB <- ts(births$TLB, start = 1960, frequency = 1)
TFR <- ts(births$TFR, start = 1960, frequency = 1)

# Training period 1960-2012 only
TLB_train <- window(TLB, start = 1960, end = 2012)
TFR_train <- window(TFR, start = 1960, end = 2012)

```




## 3. Summary Statistics
```{r summary}

# Summary statistics provide a basic numerical description of each series before visual analysis.
# We look at the minimum, maximum, mean and quartiles to understand the overall range and central tendency of TLB and TFR over the full 1960-2024 period.
cat("=== Total Live Births ===\n")
summary(TLB)

cat("\n=== Total Fertility Rate ===\n")
summary(TFR)

```
TLB ranges from a maximum of 61,775 births to a minimum of, with a mean of approximately 45,000 births per year. TFR ranges from 5.76 to a minimum of 0.97, with a mean of 2.20. Both series show a wide range between their minimum and maximum values, which is consistent with a long-run declining trend rather than a stationary series fluctuating around a constant mean.

## 4. Time Series Plots
```{r plot_tlb, fig.width=10, fig.height=5}
# Plotting the full TLB time series from 1960 to 2024.
plot(TLB,
     main = "Total Live Births in Singapore (1960-2024)",
     ylab = "Total Live Births",
     xlab = "Year")
```
```{r plot_tfr, fig.width=10, fig.height=5}
plot(TFR,
     main = "Total Fertility Rate in Singapore (1960-2024)",
     ylab = "Total Fertility Rate (per female)",
     xlab = "Year")
```
Both TLB and TFR show a clear and sustained downward trend over the entire 65-year period with no seasonality present, confirming both series are non-stationary. There are visible changes in the rate of decline at certain points in time and a brief recovery in TLB in the late 1980s, which warrant further investigation.

# Literature analysis
Literature was explored to gain an understanding of the external factors that could potentially explain the features observed in the plots. The acceleration in TLB and TFR decline visible after 1969 is consistent with the introduction of Singapore's Stop-at-Two policy, which penalised families for having more than two children and directly suppressed
birth rates through the 1970s. The brief recovery visible in TLB around 1988-1990 aligns with the government's pro-natalist reversal in 1987, which introduced financial incentives for larger families, though this was insufficient to reverse the long-run downward trajectory.These structural breaks suggest that the underlying drivers of Singapore's fertility decline are not purely demographic. They are shaped by deliberate policy decisions and external shocks that a standard statistical model alone may not fully capture.

## 6. Updated Time Series Plots

The plots are now reproduced with reference lines added at the key policy events identified above, allowing the features of the series to be interpreted more precisely.

```{r plot_tlb_annotated, fig.width=10, fig.height=5}
# Reproduce the TLB plot with policy reference lines added at the years where known government interventions occurred

plot(TLB,
     main = "Total Live Births in Singapore (1960-2024)",
     ylab = "Total Live Births",
     xlab = "Year")
# Adding lines to the time series
abline(v = 1969, col = "red",       lty = 2) 
abline(v = 1987, col = "darkgreen", lty = 2)
abline(v = 2020, col = "orange",    lty = 2)

legend("topright",
       legend = c("Stop-at-Two policy (1969)",
                  "Pro-natalist policy (1987)",
                  "COVID-19 (2020)"),
       col    = c("red", "darkgreen", "orange"),
       lty    = 2,
       cex    = 0.8)
```
```{r plot_tfr_annotated, fig.width=10, fig.height=5}
# Reproduce the TFR plot with policy reference lines and the replacement level line

plot(TFR,
     main = "Total Fertility Rate in Singapore (1960-2024)",
     ylab = "Total Fertility Rate (per female)",
     xlab = "Year")

abline(v = 1969, col = "red",       lty = 2)
abline(v = 1987, col = "darkgreen", lty = 2)
abline(v = 2020, col = "orange",    lty = 2)
abline(h = 2.1,  col = "purple",    lty = 3)

legend("topright",
       legend = c("Stop-at-Two policy (1969)",
                  "Pro-natalist policy (1987)",
                  "COVID-19 (2020)",
                  "Replacement level (2.1)"),
       col    = c("red", "darkgreen", "orange", "purple"),
       lty    = c(2, 2, 2, 3),
       cex    = 0.8)
```
With the reference lines added, the features identified in the raw plots can now be directly linked to their likely causes. The acceleration in TLB and TFR decline after 1969 aligns precisely with the Stop-at-Two policy, the brief recovery in TLB around 1988-1990 corresponds to the pro-natalist reversal, and the sharp dip in 2020 coincides with the COVID-19 pandemic. These annotated plots will serve as the basis for the temporal feature analysis that follows.

## Literature on forecasting births

To gain an understanding of how others have modelled and forecasted birth rates, a review of recent literature was carried out. He (2023) applied ARIMA models to 74 years of Chinese birth rate data and found ARIMA(0,1,2) to be the best fitting model based on AIC, with first-order differencing sufficient to achieve stationarity. Similarly, a study of Philippine birth data (European Journal of IT and Computer Science, 2025) found that ARIMA-class models outperformed other methods for forecasting declining annual birth rates in a Southeast Asian context directly comparable to Singapore.

From this literature, it can be gathered that ARIMA models with d = 1 are a reasonable starting point for data of this kind, with the AR and MA orders to be identified from the ACF and PACF of the differenced series. However, as noted in the course notes, standard ARIMA models may struggle to capture structural breaks caused by external interventions. Singapore's Stop-at-Two policy (1969) and pro-natalist reversal (1987) represent exactly this kind of disruption, meaning the final model choice may need to be justified on contextual grounds as well as purely statistical ones.

## 5. Temporal Feature Analysis

### 5.1 ACF and PACF of Raw Series
```{r acf_tlb, fig.width=10, fig.height=4}
# The ACF (correlogram) plots the autocorrelation at each lag. As noted in the course notes, if there is a trend the ACF will decrease slowly and remain significant across many lags, indicating non-stationarity.

acf(TLB, lag.max = 20,
    main = "ACF of Total Live Births (1960-2024)")
```
```{r pacf_tlb, fig.width=10, fig.height=4}
# The PACF plots the partial autocorrelation at each lag — the correlation between x_t and x_{t-k} after removing the effect of all intermediate lags.A dominant spike at lag 1 in the PACF suggests an AR(1) component may be present once the trend is removed.

pacf(TLB, lag.max = 20,
     main = "PACF of Total Live Births (1960-2024)")
```
```{r acf_tfr, fig.width=10, fig.height=4}
# ACF of TFR — we expect a similar slowly decaying pattern to TLB given both series share a strong downward trend

acf(TFR, lag.max = 20,
    main = "ACF of Total Fertility Rate (1960-2024)")
```
```{r pacf_tfr, fig.width=10, fig.height=4}
# PACF of TFR to identify any autoregressive structure that may be present in the series

pacf(TFR, lag.max = 20,
     main = "PACF of Total Fertility Rate (1960-2024)")
```

The ACF for both TLB and TFR decreases very slowly and remains significant across many lags which is a signature of a non-stationary series with a strong trend, consistent with the course notes. The PACF for both series shows a dominant spike at lag 1, suggesting that once the trend is removed an autoregressive component may be present. Differencing is required before fitting any ARIMA model.

### 5.2 First-Order Differencing

Since both series have a clear downward trend we apply first-order differencing, defined as:

∇x_t = x_t − x_{t-1}

This removes the trend and is used to determine the differencing order d in our ARIMA(p,d,q) models.
```{r diff_tlb, fig.width=10, fig.height=4}
# Apply first-order differencing to TLB training data using the diff() function and plot the result.A horizontal line at zero is added if the differenced series fluctuates around zero with no clear trend, stationarity has been achieved.

plot(diff(TLB_train),
     main = "First Difference of TLB (1960-2012)",
     ylab = "Change in TLB",
     xlab = "Year",
     col  = "steelblue",
     lwd  = 2)
abline(h = 0, lty = 2)
```
```{r diff_tfr, fig.width=10, fig.height=4}
# Apply first-order differencing to TFR training data and plot the result

plot(diff(TFR_train),
     main = "First Difference of TFR (1960-2012)",
     ylab = "Change in TFR",
     xlab = "Year",
     col  = "darkred",
     lwd  = 2)
abline(h = 0, lty = 2)
```
```{r acf_diff_tlb, fig.width=10, fig.height=4}
# ACF of the differenced TLB series.If differencing has achieved stationarity, the ACF should no longer show the slow decay pattern seen in the raw series. Significant spikes indicate the MA order q to include in the ARIMA model.

acf(diff(TLB_train), lag.max = 20,
    main = "ACF of First Differenced TLB")
```
```{r pacf_diff_tlb, fig.width=10, fig.height=4}
# PACF of the differenced TLB series.Significant spikes in the PACF indicate the AR order p to include in the ARIMA model.

pacf(diff(TLB_train), lag.max = 20,
     main = "PACF of First Differenced TLB")
```
```{r acf_diff_tfr, fig.width=10, fig.height=4}
# ACF of the differenced TFR series

acf(diff(TFR_train), lag.max = 20,
    main = "ACF of First Differenced TFR")
```
```{r pacf_diff_tfr, fig.width=10, fig.height=4}
# PACF of the differenced TFR series

pacf(diff(TFR_train), lag.max = 20,
     main = "PACF of First Differenced TFR")
```

After first-order differencing both series fluctuate around zero with no clear trend, suggesting approximate stationarity has been achieved. This indicates we should set d = 1 in our ARIMA(p,1,q) models. The ACF and PACF of the differenced series are now examined to identify suitable AR and MA orders for model fitting.

---

### 5.3 Candidate ARIMA Models (Training Period: 1960-2012)

Based on the ACF and PACF patterns of the differenced series, we fit several candidate ARIMA(p,1,q) models to the training period 1960-2012 and compare using AIC. A lower AIC indicates a better balance between model fit and complexity.
```{r models}
# We fit four candidate ARIMA(p,1,q) models to the training period 1960-2012 for each series.
# d = 1 throughout as first-order differencing was shown to achieve stationarity above. The arima() function is from base R and takes the time series and order = c(p, d, q) as arguments.

# TLB candidate models
tlb_m1 <- arima(TLB_train, order = c(1,1,0)) # AR(1) on differenced series
tlb_m2 <- arima(TLB_train, order = c(0,1,1)) # MA(1) on differenced series
tlb_m3 <- arima(TLB_train, order = c(1,1,1)) # ARMA(1,1) on differenced series
tlb_m4 <- arima(TLB_train, order = c(2,1,0)) # AR(2) on differenced series

# Compare AIC values for TLB — lower AIC is better AIC penalises model complexity so it balances fit against the number of parameters estimated
cat("=== TLB Model AIC Comparison ===\n")
cat("ARIMA(1,1,0):", AIC(tlb_m1), "\n")
cat("ARIMA(0,1,1):", AIC(tlb_m2), "\n")
cat("ARIMA(1,1,1):", AIC(tlb_m3), "\n")
cat("ARIMA(2,1,0):", AIC(tlb_m4), "\n")

# TFR candidate models — same four specifications
tfr_m1 <- arima(TFR_train, order = c(1,1,0))
tfr_m2 <- arima(TFR_train, order = c(0,1,1))
tfr_m3 <- arima(TFR_train, order = c(1,1,1))
tfr_m4 <- arima(TFR_train, order = c(2,1,0))

# Compare AIC values for TFR
cat("\n=== TFR Model AIC Comparison ===\n")
cat("ARIMA(1,1,0):", AIC(tfr_m1), "\n")
cat("ARIMA(0,1,1):", AIC(tfr_m2), "\n")
cat("ARIMA(1,1,1):", AIC(tfr_m3), "\n")
cat("ARIMA(2,1,0):", AIC(tfr_m4), "\n")

#updated residuals
tlb_best <- tlb_m2  # ARIMA(0,1,1) — lowest AIC for TLB
tfr_best <- tfr_m3  # ARIMA(1,1,1) — lowest AIC for TFR
```
For TLB, the ARIMA(0,1,1) model produced the lowest AIC of 975.79, making it the preferred candidate for the training period 1960-2012. The ARIMA(1,1,0) model was a close second with an AIC of 975.81, suggesting that both a moving average and autoregressive specification of order 1 describe the differenced TLB series similarly well. The more complex models ARIMA(1,1,1) and ARIMA(2,1,0) both produced higher AIC values, indicating the added parameters did not
sufficiently improve the fit to justify their inclusion.

For TFR, the ARIMA(1,1,1) model produced the lowest AIC of -44.77 and is selected as the preferred candidate. Note that for TFR the AIC values are negative — this is not unusual and simply reflects the scale of the data, as TFR values are small decimals compared to the large counts in TLB. A more negative AIC still indicates a better fitting model. The ARIMA(1,1,0) and ARIMA(2,1,0) models were comparable to each other at around -39, while ARIMA(0,1,1) performed notably worse at -34.58, suggesting that an autoregressive component is important for capturing the structure of the differenced TFR series.

Based on these results, `tlb_best` is updated to `tlb_m2` (ARIMA(0,1,1)) and `tfr_best` is updated to `tfr_m3` (ARIMA(1,1,1)) for the residual diagnostics and forecasting that follow.

---

### 5.4 Residual Diagnostics
A well-fitting model should produce residuals resembling white noise, so no remaining autocorrelation structure. If the model has captured all the systematic information in the data, what remains should be purely random.
```{r residuals}
# Select the best model for each series based on the lowest AIC from Section 5.3 above.
#Update tlb_m1 and tfr_m1 to whichever model had the lowest AIC — e.g. change to tlb_m2 if ARIMA(0,1,1) produced the lowest AIC for TLB.

tlb_best <- tlb_m2  # ARIMA(0,1,1) - lowest AIC for TLB
tfr_best <- tfr_m3  # ARIMA(1,1,1) - lowest AIC for TFR

# Print the full model output including parameter estimates and standard errors
cat("=== Best TLB Model ===\n")
print(tlb_best)

cat("\n=== Best TFR Model ===\n")
print(tfr_best)
```
```{r resid_tlb, fig.width=10, fig.height=4}
# ACF of the TLB model residuals.
# As noted in the course notee, if the residuals resemble white noise the ACF  should show no significant spikes beyond the dotted blue significance bounds.
# na.omit() removes the NA introduced by differencing.

acf(na.omit(tlb_best$residuals), lag.max = 20,
    main = "ACF of Residuals: Best TLB Model")
```
```{r resid_pacf_tlb, fig.width=10, fig.height=4}
# PACF of the TLB model residuals.Any significant spikes here would suggest a remaining AR structure that has not been captured by the model.

pacf(na.omit(tlb_best$residuals), lag.max = 20,
     main = "PACF of Residuals: Best TLB Model")
```
```{r resid_tfr, fig.width=10, fig.height=4}
# ACF of the TFR model residuals

acf(na.omit(tfr_best$residuals), lag.max = 20,
    main = "ACF of Residuals: Best TFR Model")
```
```{r resid_pacf_tfr, fig.width=10, fig.height=4}
# PACF of the TFR model residuals

pacf(na.omit(tfr_best$residuals), lag.max = 20,
     main = "PACF of Residuals: Best TFR Model")
```
```{r qqplots, fig.width=10, fig.height=4}
# Q-Q plots check whether the residuals are approximately normally distributed, a key assumption of the ARIMA model.Points close to the diagonal red line indicate the normality assumption is reasonable. par(mfrow) sets up a side-by-side plot layout.

par(mfrow = c(1, 2))

qqnorm(na.omit(tlb_best$residuals),
       main = "Q-Q Plot: TLB Residuals")
qqline(na.omit(tlb_best$residuals), col = "red")

qqnorm(na.omit(tfr_best$residuals),
       main = "Q-Q Plot: TFR Residuals")
qqline(na.omit(tfr_best$residuals), col = "red")

```

Three diagnostic tools are used to assess model adequacy. The ACF and PACF of the residuals check whether any autocorrelation structure remains after fitting, the residuals should resemble white noise with no significant spikes. The Q-Q plot checks whether the residuals are approximately normally distributed, which is a key assumption of the ARIMA model.

For the TLB ARIMA(0,1,1) model, the ACF of residuals shows all spikes within the significance bounds with only a marginal spike around lag 12, and the PACF is similarly clean. This is consistent with white noise, suggesting the model has adequately captured the structure of the data. The Q-Q plot shows points following the diagonal line closely through the central range with slight deviation at the upper tail, indicating mild but acceptable departures 
from normality.

For the TFR ARIMA(1,1,1) model, both the ACF and PACF of residuals show no significant spikes, which is a strong indication of white noise behaviour and a good model fit. The Q-Q plot shows reasonable alignment with the diagonal line, with some deviation at the lower tail suggesting a slight left skew. This may reflect the occasional large 
drops in TFR caused by Singapore's policy interventions, which the model cannot fully account for. Overall both models appear adequate for preliminary forecasting purposes.

---

### 5.5 Preliminary Forecast (2013-2024)

Using the best fitting models we forecast TLB and TFR for 2013-2024 and compare against actual observed values.The vertical dotted line marks the end of the training period at 2012 — everything to the right is forecast territory. The shaded region represents the 95% prediction interval, within which we expect the true values to fall if the model is adequate.
```{r forecast, fig.width=10, fig.height=5}
# Use the predict() function to forecast 12 years ahead from the end of the training period (2012) to 2024.
# predict() returns pred (the forecast values) and se (the standard errors used to compute intervals).
tlb_fc <- predict(tlb_best, n.ahead = 12)
tfr_fc <- predict(tfr_best, n.ahead = 12)


plot(TLB,
     main = "TLB: Observed vs Forecast (2013-2024)",
     ylab = "Total Live Births",
     xlab = "Year",
     col  = "steelblue",
     lwd  = 2)

# Overlay the forecast values in red
lines(tlb_fc$pred, col = "red", lwd = 2)

# Add 95% prediction intervals: forecast +/- 1.96 * SE
# 1.96 corresponds to the 97.5th percentile of the 
# standard normal distribution
lines(tlb_fc$pred + 1.96 * tlb_fc$se, col = "red", lty = 2)
lines(tlb_fc$pred - 1.96 * tlb_fc$se, col = "red", lty = 2)

# Vertical dotted line marking end of training period
abline(v = 2012, lty = 3, col = "grey40")

legend("topright",
       legend = c("Observed", "Forecast", "95% PI"),
       col    = c("steelblue", "red", "red"),
       lty    = c(1, 1, 2),
       cex    = 0.8)
```

```{r forecast_tfr, fig.width=10, fig.height=5}
# TFR forecast plot — same structure as TLB above
plot(TFR,
     main = "TFR: Observed vs Forecast (2013-2024)",
     ylab = "Total Fertility Rate",
     xlab = "Year",
     col  = "darkred",
     lwd  = 2)

# Overlay TFR forecast and prediction intervals in blue
lines(tfr_fc$pred, col = "blue", lwd = 2)
lines(tfr_fc$pred + 1.96 * tfr_fc$se, col = "blue", lty = 2)
lines(tfr_fc$pred - 1.96 * tfr_fc$se, col = "blue", lty = 2)

# Vertical dotted line marking end of training period
abline(v = 2012, lty = 3, col = "grey40")

legend("topright",
       legend = c("Observed", "Forecast", "95% PI"),
       col    = c("darkred", "blue", "blue"),
       lty    = c(1, 1, 2),
       cex    = 0.8)
```

The ARIMA(0,1,1) model forecasts TLB remaining approximately constant at around 43,000 births per year from 2013 onwards. The actual observed values however continue declining to around 33,000 by 2024, falling well below the forecast. While the observed values remain within the 95% prediction interval, this is largely due to the 
interval widening considerably over time rather than the model accurately predicting the direction of the series. Overall the model underestimates the continued decline in TLB, suggesting that alternative specifications should be explored.

The ARIMA(1,1,1) model forecasts TFR remaining approximately constant at around 1.2 from 2013 onwards. The actual observed values track very closely to the forecast through most of the period, declining slightly from around 1.2 to 
0.97 by 2024. The observed values remain well within the 95% prediction interval throughout, and the forecast direction is broadly consistent with reality. Overall the ARIMA(1,1,1) model performs considerably better for TFR than the ARIMA(0,1,1) does for TLB, suggesting it is a reasonable candidate model to carry forward into the Final Report.
---

## 6. Research Questions

Based on the EDA the following research questions are proposed for the Final Report:

**RQ1:** Can ARIMA models fitted to Singapore's TLB and TFR data from 1960-2012 produce accurate forecasts for
2013-2024, and how does accuracy compare across different model specifications?

**RQ2:** Does the TFR series require a different model structure than TLB, given that TFR is a rate measure and may be affected differently by policy-driven structural breaks?

**RQ3:** To what extent do social and economic factors such as the Stop-at-Two policy (1969) and pro-natalist reversal (1987) limit the predictive power of standard ARIMA models for Singapore's birth data?

**Proposed analyses for Part 2:**

- Fit at least 2 ARIMA models per series on 1960-2012
- Compare models using AIC and residual diagnostics
- Forecast 2013-2024 and compare against actual values
- Consider whether structural breaks warrant subsetting

**Anticipated limitations:**

- Only 53 training observations limits model complexity
- Policy-driven structural breaks may not be captured by standard ARIMA models
- TLB depends on both fertility rates and population size which may confound modelling
- The TFR definition change at 1980 introduces a subtle discontinuity

---

## 7. Statistical Appendix

### 7.1 ARIMA(p, d, q) Model

An ARIMA(p, d, q) model applied to a time series x_t is defined as:

φ(B) ∇^d x_t = θ(B) w_t

where:

- B is the backshift operator: B x_t = x_{t-1}
- ∇^d = (1 − B)^d is the differencing operator of order d
- φ(B) = 1 − φ_1 B − ... − φ_p B^p is the AR polynomial
- θ(B) = 1 + θ_1 B + ... + θ_q B^q is the MA polynomial
- w_t ~ N(0, σ²) is white noise

### 7.2 Autocorrelation Function (ACF)

The sample autocorrelation at lag k is:

r_k = Σ(x_t − x̄)(x_{t−k} − x̄) / Σ(x_t − x̄)²

Under H₀: ρ_k = 0, significance bounds are plotted at ±2/√n. Any spike beyond these bounds indicates significant autocorrelation at that lag.

### 7.3 Partial Autocorrelation Function (PACF)

The partial autocorrelation at lag k measures the correlation between x_t and x_{t−k} after removing the linear effects of all intermediate observations. A pure AR(p) process has PACF that cuts off after lag p.

### 7.4 Akaike Information Criterion (AIC)

AIC = −2 log(L) + 2k

where L is the maximised likelihood and k is the number of estimated parameters. Lower AIC indicates a better trade-off between model fit.

### 7.5 White Noise

A process w_t is white noise if:

- E[w_t] = 0 for all t
- Var(w_t) = σ² for all t
- Cov(w_t, w_s) = 0 for all t ≠ s

Residuals from a well-fitted model should behave as white noise, confirmed by inspecting the ACF and PACF of residuals for any remaining significant spikes.

---

## References

He, Y. (2023). Forecasts and Analyzes of China's Birth Rate. Proceedings of DSIT 2023. 
https://www.scitepress.org/Papers/2023/128198/

European Journal of Information Technologies and Computer Science (2025). Forecasting Number of Births 
in the Philippines Using SARIMA.https://eu-opensci.org/index.php/compute/article/view/10129

United Nations, Department of Economic and Social Affairs, Population Division (2023). World Population Prospects 2023. Retrieved from https://population.un.org/wpp/

Singapore Department of Statistics (2026). Births and Fertility Rates, Annual. Retrieved from 
https://data.gov.sg
