---
title: "Forecasting Singapore's Total Live Births and Total Fertility Rate: The Role of the Chinese Zodiac Dragon Year Effect"
author: "Kiran Shankaran"
date: "8 June 2026"
output:
  pdf_document:
    number_sections: false
    fig_caption: true
geometry: margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \onehalfspacing
---

```{r setup, include=FALSE}
# Global chunk options: hide code from PDF body, hide messages and 
# warnings, set default figure dimensions and centre alignment.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, fig.align = "center")
```

```{r load-data, include=FALSE}
# Read the raw SingStat Table M810091 export. The file is in wide format 
# with years as columns running from 2025 (left) to 1960 (right). Missing 
# values are encoded as the string "na" so we declare that on read.

raw <- read.csv("BirthsAndFertilityRates.csv",
                na.strings = "na",
                stringsAsFactors = FALSE,
                check.names = FALSE)

# Some row labels in the source CSV are indented with leading spaces 
# (e.g. "    Chinese"). Strip the whitespace so each variable name 
# matches cleanly when we extract rows.
raw$DataSeries <- trimws(raw$DataSeries)

# Helper function: pull a named row from the raw data, convert to 
# numeric, reverse year order so 1960 comes first, and drop the 
# incomplete 2025 column. Returns a length-65 numeric vector.
extract_row <- function(name) {
  row_vals <- raw[raw$DataSeries == name, -1]
  values <- as.numeric(unlist(row_vals))
  values <- rev(values)
  values[1:65]
}

# Build a single tidy dataframe with one row per year and one column 
# per variable. R-friendly names replace the original source labels.
fertility <- data.frame(
  Year         = 1960:2024,
  TFR          = extract_row("Total Fertility Rate (TFR)"),
  TLB          = extract_row("Total Live-Births"),
  TFR_Chinese  = extract_row("Chinese"),
  TFR_Malay    = extract_row("Malays"),
  TFR_Indian   = extract_row("Indians"),
  TFR_15_19    = extract_row("15 - 19 Years"),
  TFR_20_24    = extract_row("20 - 24 Years"),
  TFR_25_29    = extract_row("25 - 29 Years"),
  TFR_30_34    = extract_row("30 - 34 Years"),
  TFR_35_39    = extract_row("35 - 39 Years"),
  TFR_40_44    = extract_row("40 - 44 Years"),
  TFR_45_49    = extract_row("45 - 49 Years"),
  CBR          = extract_row("Crude Birth Rate"),
  GRR          = extract_row("Gross Reproduction Rate"),
  NRR          = extract_row("Net Reproduction Rate")
)

# Convert variables used in modelling to ts objects. Base R's arima() 
# and related time series functions require ts inputs rather than 
# dataframe columns.
TLB <- ts(fertility$TLB,         start = 1960, frequency = 1)
TFR <- ts(fertility$TFR,         start = 1960, frequency = 1)
TFR_chinese <- ts(fertility$TFR_Chinese, start = 1960, frequency = 1)
TFR_malay   <- ts(fertility$TFR_Malay,   start = 1960, frequency = 1)
TFR_indian  <- ts(fertility$TFR_Indian,  start = 1960, frequency = 1)

# Training period for model estimation runs from 1960 to 2012. The 
# remaining years (2013 to 2024) form the test set used to evaluate 
# forecast accuracy.
TLB_train <- window(TLB, end = 2012)
TFR_train <- window(TFR, end = 2012)
```

# Abstract



GitHub repository: <https://github.com/kiranshank12519/Time-Series-Project-Singapore-Births-Forecast>

---

# 1. Introduction

This study evaluates whether incorporating a 12-year seasonal component, 
reflecting the documented Chinese zodiac Dragon Year fertility effect 
(Agarwal et al., 2021), improves forecast accuracy for Singapore's Total 
Live Births (TLB) and Total Fertility Rate (TFR) compared to standard 
non-seasonal ARIMA models.

The motivation arises from the exploratory analysis conducted in Part 1, 
in which an ARIMA(0,1,1) model was selected as the best non-seasonal fit 
for TLB but produced a flat forecast that systematically underestimated 
the continued decline observed between 2013 and 2024. A marginal spike at 
lag 12 was visible in the residual autocorrelation function but not 
investigated further. Singapore's demographic literature offers a clear 
hypothesis for this periodicity: years aligned with the Chinese zodiac 
Dragon experience elevated birth rates within the Chinese-majority 
population, producing a 12-year cycle that overlays the long-run 
downward trend.

*[Remainder of Introduction to be drafted Day 3.]*

# 2. Literature Review

*[Placeholder — Day 3. Will cover: Singapore fertility policy timeline 
(anti-natalist 1960s-80s, pro-natalist from 1987, M&P Packages 
2001-2015); documented Chinese zodiac Dragon Year fertility effect 
(Agarwal et al., 2021); prior ARIMA and SARIMA applications to fertility 
forecasting in Singapore and comparable Asian contexts.]*

# 3. Methods

## 3.1 Data

The data were sourced from the Singapore Department of Statistics 
(SingStat Table M810091), covering annual observations from 1960 to 2024. 
The variables used in this analysis are Total Live Births (TLB), the 
aggregate Total Fertility Rate (TFR), and TFR disaggregated by ethnic 
group (Chinese, Malay, and Indian). Standard preparation was applied: 
year order was reversed to chronological, missing values encoded as the 
string "na" in the source file were converted to `NA`, and the incomplete 
2025 column was excluded. All series were converted to time series 
objects with annual frequency for subsequent modelling.

## 3.2 Stationarity Testing


## 3.3 Non-seasonal ARIMA Models



## 3.4 Seasonal ARIMA Models



## 3.5 Residual Diagnostics



## 3.6 Forecast Evaluation


# 4. Results



# 5. Discussion



# 6. Conclusion



# References



\newpage

# Statistical Appendix

## A.1 ARIMA(p, d, q) Model

An ARIMA(p, d, q) model applied to a time series $x_t$ is defined as:

$$\phi(B) \, \nabla^d x_t = \theta(B) \, w_t$$

where $B$ is the backshift operator ($B x_t = x_{t-1}$), 
$\nabla^d = (1 - B)^d$ is the differencing operator of order $d$, 
$\phi(B) = 1 - \phi_1 B - \cdots - \phi_p B^p$ is the autoregressive 
polynomial, $\theta(B) = 1 + \theta_1 B + \cdots + \theta_q B^q$ is the 
moving average polynomial, and $w_t \sim N(0, \sigma^2)$ is white noise.

## A.2 Seasonal ARIMA Model

The seasonal extension SARIMA$(p, d, q)(P, D, Q)_s$ adds seasonal AR, 
differencing, and MA terms at lag $s$:

$$\Phi(B^s) \, \phi(B) \, \nabla_s^D \, \nabla^d x_t = \Theta(B^s) \, \theta(B) \, w_t$$

where $\Phi$ and $\Theta$ are the seasonal AR and MA polynomials in $B^s$, 
and $\nabla_s = 1 - B^s$ is the seasonal differencing operator. For this 
study, $s = 12$ to reflect the proposed Dragon Year periodicity.

## A.3 Autocorrelation and Partial Autocorrelation Functions

The sample autocorrelation at lag $k$ is

$$r_k = \frac{\sum_{t=k+1}^{n} (x_t - \bar{x})(x_{t-k} - \bar{x})}{\sum_{t=1}^{n} (x_t - \bar{x})^2}$$

Significance bounds under $H_0: \rho_k = 0$ are plotted at $\pm 2/\sqrt{n}$. 
The partial autocorrelation at lag $k$ measures the correlation between 
$x_t$ and $x_{t-k}$ after removing the linear effects of intermediate 
observations.

## A.4 KPSS Test for Stationarity

The Kwiatkowski–Phillips–Schmidt–Shin test takes the null hypothesis that 
the series is (level- or trend-) stationary. A rejection of $H_0$ at the 
chosen significance level indicates non-stationarity requiring 
transformation or differencing.

## A.5 Ljung-Box Test for White Noise Residuals

The Ljung-Box statistic tests the joint null that the first $h$ 
autocorrelations of the residuals are zero:

$$Q = n(n+2) \sum_{k=1}^{h} \frac{\hat{r}_k^2}{n-k}$$

Under $H_0$, $Q$ follows a $\chi^2$ distribution with $h - p - q$ 
degrees of freedom for an ARIMA$(p, d, q)$ model. Failure to reject 
$H_0$ provides evidence that residuals behave as white noise.

## A.6 Akaike Information Criterion

$$\mathrm{AIC} = -2 \log L + 2k$$

where $L$ is the maximised likelihood and $k$ is the number of estimated 
parameters. Lower AIC values indicate a better trade-off between model 
fit and complexity.

## A.7 Forecast Accuracy Metrics

Root Mean Squared Error and Mean Absolute Percentage Error are used to 
compare forecast performance on the test period (2013-2024):

$$\mathrm{RMSE} = \sqrt{\frac{1}{h} \sum_{t=1}^{h} (y_t - \hat{y}_t)^2}
\qquad
\mathrm{MAPE} = \frac{100}{h} \sum_{t=1}^{h} \left| \frac{y_t - \hat{y}_t}{y_t} \right|$$

where $y_t$ is the observed value and $\hat{y}_t$ is the forecast at 
horizon $t$, with $h = 12$ here.